---
title: "Imputations and data cleaning Lathyrus 2006-2017"
output:
  pdf_document: default
---

```{r include=FALSE}
library(tidyr)
library(plyr)
library(sjPlot)
library(ggplot2)
library(fmsb)
library(xtable)
library(gridExtra)
library(MASS)
options(xtable.floating = FALSE)
options(xtable.timestamp = "")
options(xtable.comment = FALSE)

data_imput<-read.table("C:/Users/avald/Dropbox/SU/Projects/lathyrus/lathyrus_ms1/data/clean/data_20062017_forimput.csv",
                       header=T,sep="\t",dec=",") #Read all data
bud_sizes<-read.table("C:/Users/avald/Dropbox/SU/Projects/lathyrus/lathyrus_ms1/data/clean/bud_sizes.csv",
                       header=T,sep="\t",dec=",") #Read data on bud sizes
head(data_imput)
str(data_imput)
data_imput$FFD_corr<-as.Date(as.character(data_imput$FFD_corr),format="%d/%m/%Y")
data_imput$vernal<-as.POSIXct(data_imput$vernal,format="%d/%m/%y %H:%M")
data_imput$FFD<-as.numeric(with(data_imput,as.POSIXct(FFD_corr)-vernal))

subset(data_imput,n_fr>cum_n_fl) #4 cases where n_fruits>n_flowers --> fix
#Equal n_fl to n_fr
data_imput$cum_n_fl<-with(data_imput,ifelse(n_fr>cum_n_fl,n_fr,cum_n_fl))

head(bud_sizes)

bud_sizes[sapply(bud_sizes, function(x) all(is.na(x)))] <- NULL #remove columns with all NAs
str(bud_sizes)
bud_sizes<-gather(bud_sizes, key=day, value=bud_size, X114:X148,factor_key=T) #Convert from wide to long
bud_sizes<-subset(bud_sizes,!bud_size=="") #remove rows with blank bud_size
bud_sizes$day<-as.integer(substring(as.character(bud_sizes$day),2))
bud_sizes$bud_size[bud_sizes$bud_size=="S"] <- 1
bud_sizes$bud_size[bud_sizes$bud_size=="M"] <- 2
bud_sizes$bud_size[bud_sizes$bud_size=="L"] <- 3
bud_sizes$bud_size[bud_sizes$bud_size=="XL"] <-4
bud_sizes$bud_size<-as.integer(bud_sizes$bud_size)

head(bud_sizes)
str(bud_sizes)

#subset to use for building a model of FFD against bud size and day
subset_model<-subset(bud_sizes,!FFD_action=="impute")

with(subset_model,hist(FFD_julian,breaks=50))
with(subset_model,hist(bud_size))
with(subset_model,hist(day))
with(subset_model,plot(as.factor(bud_size)~day))
with(subset_model,plot(FFD_julian~as.factor(bud_size)))
with(subset_model,plot(FFD_julian~day))
```

I used information on buds sizes to impute FFD. The day when the bud was observed was also included in the model, as well as the interaction, to account for the fact that plants might develop faster/slower in the beginning/end of the season. First, I fitted a model with all years (but note that 2016 and 2017 had no information on bud sizes at all).


```{r echo=FALSE, fig.height=4, fig.width=4}
model_FFD<-lm(FFD_julian~bud_size*day,subset_model)
summary(model_FFD)
par(mfrow=c(2,2)) 
plot(model_FFD)

```

Then I fitted a different model for each year

```{r}
model_FFD06<-lm(FFD_julian~bud_size+day,subset(subset_model,year==2006))
model_FFD07<-lm(FFD_julian~bud_size+day,subset(subset_model,year==2007)) 
model_FFD08<-lm(FFD_julian~bud_size+day,subset(subset_model,year==2008)) 
model_FFD09<-lm(FFD_julian~bud_size+day,subset(subset_model,year==2009)) 
model_FFD10<-lm(FFD_julian~bud_size*day,subset(subset_model,year==2010)) 
model_FFD11<-lm(FFD_julian~bud_size*day,subset(subset_model,year==2011)) 
model_FFD12<-lm(FFD_julian~bud_size*day,subset(subset_model,year==2012)) 
model_FFD13<-lm(FFD_julian~bud_size*day,subset(subset_model,year==2013)) 
model_FFD14<-lm(FFD_julian~bud_size*day,subset(subset_model,year==2014)) 
model_FFD15<-lm(FFD_julian~bud_size+day,subset(subset_model,year==2015)) 

summary(model_FFD06)
summary(model_FFD07)
summary(model_FFD08)
summary(model_FFD09)
summary(model_FFD10)
summary(model_FFD11)
summary(model_FFD12)
summary(model_FFD13)
summary(model_FFD14)
summary(model_FFD15)
```

Compare distributions of observed vs predicted values

```{r echo=FALSE}
data06<-data.frame(observed=subset(subset_model,year==2006)$FFD_julian, predicted=predict(model_FFD06))
data07<-data.frame(observed=subset(subset_model,year==2007)$FFD_julian, predicted=predict(model_FFD07))
data08<-data.frame(observed=subset(subset_model,year==2008)$FFD_julian, predicted=predict(model_FFD08))
data09<-data.frame(observed=subset(subset_model,year==2009)$FFD_julian, predicted=predict(model_FFD09))
data10<-data.frame(observed=subset(subset_model,year==2010)$FFD_julian, predicted=predict(model_FFD10))
data11<-data.frame(observed=subset(subset_model,year==2011)$FFD_julian, predicted=predict(model_FFD11))
data12<-data.frame(observed=subset(subset_model,year==2012)$FFD_julian, predicted=predict(model_FFD12))
data13<-data.frame(observed=subset(subset_model,year==2013)$FFD_julian, predicted=predict(model_FFD13))
data14<-data.frame(observed=subset(subset_model,year==2014)$FFD_julian, predicted=predict(model_FFD14))
data15<-data.frame(observed=subset(subset_model,year==2015)$FFD_julian, predicted=predict(model_FFD15))

grid.arrange(
ggplot(data06) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2006"),
ggplot(data07) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2007"),
ggplot(data08) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2008"),
ggplot(data09) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2009"),
ggplot(data10) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2010"),
ggplot(data11) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2011"),
ggplot(data12) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2012"),
ggplot(data13) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2013"),
ggplot(data14) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2014"),
ggplot(data15) +geom_density(aes(x=observed),alpha=.3, fill="blue")+geom_density(aes(x=predicted),alpha=.3, fill="red")+ggtitle("2015"),
ncol=3)
```

```{r include=FALSE}
#Use per-year models 
#to predict FFD values where FFD_action=biased/imp_bdsize/impute

predict06<-as.data.frame(predict(model_FFD06,newdata=subset(subset(bud_sizes,year==2006),FFD_action=="impute")))
predict07<-as.data.frame(predict(model_FFD07,newdata=subset(subset(bud_sizes,year==2007),FFD_action=="impute")))
predict08<-as.data.frame(predict(model_FFD08,newdata=subset(subset(bud_sizes,year==2008),FFD_action=="impute")))
predict09<-as.data.frame(predict(model_FFD09,newdata=subset(subset(bud_sizes,year==2009),FFD_action=="impute")))
predict10<-as.data.frame(predict(model_FFD10,newdata=subset(subset(bud_sizes,year==2010),FFD_action=="impute")))
predict11<-as.data.frame(predict(model_FFD11,newdata=subset(subset(bud_sizes,year==2011),FFD_action=="impute")))
predict12<-as.data.frame(predict(model_FFD12,newdata=subset(subset(bud_sizes,year==2012),FFD_action=="impute")))
predict13<-as.data.frame(predict(model_FFD13,newdata=subset(subset(bud_sizes,year==2013),FFD_action=="impute")))
predict14<-as.data.frame(predict(model_FFD14,newdata=subset(subset(bud_sizes,year==2014),FFD_action=="impute")))
predict15<-as.data.frame(predict(model_FFD15,newdata=subset(subset(bud_sizes,year==2015),FFD_action=="impute")))

names(predict06)<-c("FFD_i")
names(predict07)<-c("FFD_i")
names(predict08)<-c("FFD_i")
names(predict09)<-c("FFD_i")
names(predict10)<-c("FFD_i")
names(predict11)<-c("FFD_i")
names(predict12)<-c("FFD_i")
names(predict13)<-c("FFD_i")
names(predict14)<-c("FFD_i")
names(predict15)<-c("FFD_i")

predictions<-rbind(predict06,predict07,predict08,predict09,predict10,predict11,predict12,predict13,predict14,predict15)

bud_sizes<-merge(bud_sizes, predictions, by="row.names",all.x=T) #New colum with imputed FFD (FFD_i) for pls where FFD_action=biased/imp_bdsize/impute
bud_sizes$Row.names<-NULL
bud_sizes<-unique(bud_sizes[c("year", "id","FFD_julian","FFD_action","FFD_i")])
bud_sizes$FFD_buds<-ifelse(is.na(bud_sizes$FFD_julian),bud_sizes$FFD_i,bud_sizes$FFD_julian)
bud_sizes$FFD_imputed<-ifelse(bud_sizes$FFD_action=="impute",1,0) #Says if FFD was imputed (1) or not (0)

head(bud_sizes)
str(bud_sizes)

#merge with data imput
data_imput<-merge(data_imput,bud_sizes[c(1:2,6,7)],by=c("year","id"),all.x=T,all.y=T)
data_imput$FFD<-ifelse(is.na(data_imput$FFD),data_imput$FFD_buds,data_imput$FFD)
data_imput$FFD_buds<-NULL
head(data_imput)
str(data_imput)

subset(data.frame(with(data_imput,table(year,id))),Freq>1) #No repeated ids!

```

Results of the imputation
```{r}
nrow(subset(data_imput,FFD_action=="impute"&!is.na(FFD))) #56 cases imputed
nrow(subset(data_imput,FFD_action=="impute"&is.na(FFD))) #46 cases could not be imputed because no info on bud sizes
nrow(subset(data_imput,is.na(FFD)&data==1)) #46 cases where FFD is NA = cases that could not be imputed --> mark as missing data
```

Predict shoot volume with number of flowers?

```{r}
nrow(subset(data_imput,data==1&is.na(shoot_vol))) #37 pls with no shoot_vol
nrow(subset(data_imput,data==1&is.na(cum_n_fl))) #24 pls with no cum_n_fl
nrow(subset(data_imput,data==1&is.na(cum_n_fl)&is.na(shoot_vol))) 
#3 pls with no shoot_vol and no cum_n_fl --> mark as missing data
nrow(subset(data_imput,is.na(cum_n_fl)&data==1&!is.na(shoot_vol))) 
#We can predict shoot_vol from cum_n_fl in 24-3=21 pls
```




