---
title: "Yearly selection models Lathyrus"
output:
  pdf_document: 
    toc: yes
    toc_depth: 4
editor_options:
  chunk_output_type: inline
---

```{r load image and packages, include=FALSE}
load(file = "C:/Users/User/Dropbox/SU/Projects/lathyrus/code/image1.RData")
library(dplyr)
library(broom)
library(knitr)
library(car)
library(ggplot2)
library(fmsb)
library(gridExtra)
library(papeR)
```
\newpage
# Select data and look at variables

```{r Select data and look at variables, echo=TRUE}
data_sel<-subset(alldata_weather_subs,!is.na(n_fl)&!is.na(FFD)) 
#Select data where both FFD and n_fl are available
nrow(subset(data_sel,is.na(n_intact_seeds))) #No NAs for seed data
```

```{r Histograms, echo=FALSE, fig.height=4, fig.width=12}
par(mfrow=c(1,3))
with(data_sel,hist(FFD))
with(data_sel,hist(n_fl))
with(data_sel,hist(n_intact_seeds))
```

# Calculation of relative fitness and standardized traits  
Relativization and standardization was done within each year.

```{r Relative fitness and standardized traits}
data_sel<-data.frame(
  data_sel %>% 
  group_by(year) %>% 
  mutate(n_intact_seeds_rel=n_intact_seeds/mean(n_intact_seeds)) %>% #Relative fitness
  mutate(FFD_std=(FFD-mean(FFD))/sd(FFD)) %>%                        #Standardized FFD
  mutate(n_fl_std=(n_fl-mean(n_fl))/sd(n_fl)))                       #Standardized n_fl
```

# Phenotypic selection models with all data

```{r Phenotypic selection models all data}
Anova(lm(n_intact_seeds_rel ~ FFD_std+FFD_std:year, data = data_sel),type="II") 
#Not sure about type - II for interactions?
Anova(lm(n_intact_seeds_rel ~ FFD_std+FFD_std:year+n_fl_std, 
         data = data_sel),type="II")
#Selection for early flowering differs among years when considering only FFD
#and also when including number of flowers as a covariate

Anova(update(lm(n_intact_seeds_rel ~ (FFD_std+I(FFD_std^2))*year,
                data = data_sel),.~.-year),type="II") 
Anova(update(lm(n_intact_seeds_rel~(FFD_std+I(FFD_std^2))*year+
                  FFD_std:n_fl_std+n_fl_std+I(n_fl_std^2), data = data_sel),.~.-year),type="II") 
#No evidence of non-linear selection for FFD (but quadratic selection for n_fl)
```

# Phenotypic selection models for each year

## With only FFD

```{r Selection models with FFD}
sel_models1<-data_sel %>% 
  group_by(year) %>% 
  do(model = lm(n_intact_seeds_rel ~ FFD_std, data = .)) 
sel_models1_coefs<-data.frame(sel_models1 %>% tidy(model))
sel_models1_coefs$sig<-ifelse(sel_models1_coefs$p.value<0.05,"*","") 
sel_models1_rsq<-data.frame(sel_models1 %>% glance(model))[1:3]
sel_models1_anova<-cbind(
  year=c("1987","1988","1989","1990","1991","1992","1993","1994","1995","1996","2006",
         "2007","2008","2009","2010","2011","2012","2013","2014","2015","2016","2017"),
  variable=rep(c("FFD_std"),22),
   plyr::ldply(lapply(as.list(sel_models1)$model,FUN=Anova), function(x) data.frame(x)[1,3:4]))
sel_models1_anova$sig<-ifelse(sel_models1_anova$Pr..F.<0.05,"*","")
kable(sel_models1_coefs)  #Selection gradients
#FFD * (selection for early flowering) in all years but 1995,2009,2013,2015,2017
kable(sel_models1_rsq) #R squares
kable(sel_models1_anova) #Anova results
```

Plots seeds vs FFD

```{r Plots fitness vs FFD, echo=FALSE, fig.height=3, fig.width=5}
ggplot(data_sel,aes(FFD_std,n_intact_seeds_rel,colour=year))+
  geom_smooth(method="lm",se=F,size=0.6)+geom_point(size=0.5)+
  xlab("Standardized FFD")+ylab("Relative number of intact seeds")
```

Same plot but with only the fit lines (no data points)

```{r Plots fitness vs FFD no points, echo=FALSE, fig.height=3, fig.width=5}
ggplot(data_sel,aes(FFD_std,n_intact_seeds_rel,colour=year))+
  geom_smooth(method="lm",se=F,size=0.6)+
  xlab("Standardized FFD")+ylab("Relative number of intact seeds")
```

There is one outlier in 2017 with a very high relative number of intact seeds. This is a correct value (the plant had 9 intact seeds but the relative value is so high because only 7 plants produced seeds in 2017, so the mean for that year is very low). I think this value is driving the pattern for 2017 (year with the most positive slope in the previous graph).

## With FFD & number of flowers

```{r Selection models with FFD & n_fl}
sel_models2<-as.list(data_sel %>% 
  group_by(year) %>% 
  do(model = lm(n_intact_seeds_rel ~ FFD_std+n_fl_std, data = .)) )
sel_models2<-data_sel %>% 
  group_by(year) %>% 
  do(model = lm(n_intact_seeds_rel ~ FFD_std+n_fl_std, data = .)) 
sel_models2_coefs<-data.frame(sel_models2 %>% tidy(model))
sel_models2_coefs$sig<-ifelse(sel_models2_coefs$p.value<0.05,"*","") 
sel_models2_rsq<-data.frame(sel_models2 %>% glance(model))[1:3]
sel_models2_anova<-cbind(
  year=c("1987","1987","1988","1988","1989","1989","1990","1990","1991","1991",
  "1992","1992","1993","1993","1994","1994","1995","1995","1996","1996","2006",
  "2006","2007","2007","2008","2008","2009","2009","2010","2010","2011","2011",
  "2012","2012","2013","2013","2014","2014","2015","2015","2016","2016",
  "2017","2017"),variable=rep(c("FFD_std","n_fl_std"),22),
   plyr::ldply(lapply(as.list(sel_models2)$model,FUN=Anova), 
   function(x) data.frame(x)[1:2,3:4]))
sel_models2_anova$sig<-ifelse(sel_models2_anova$Pr..F.<0.05,"*","") 

kable(sel_models2_coefs) #Selection gradients
#FFD * (selection for early flowering) in 1991,1992,1993,2007,2010,2012,2014
kable(sel_models2_rsq) #R squares
kable(sel_models2_anova) #Anova results

sel_grads_FFD<-subset(sel_models2_coefs,term=="FFD_std")[c(1,3,7)]
sel_grads_FFD 
#These are the per-year selection gradients for FFD, to be used in further analyses
```

## Non-linear selection (with FFD & number of flowers)

```{r Non-linear selection}
sel_models3<-data_sel %>% 
  group_by(year) %>% 
  do(model = lm(n_intact_seeds_rel ~ FFD_std+I(FFD_std^2)+n_fl_std+I(n_fl_std^2)+
                  FFD_std:n_fl_std, data = .)) 
sel_models3_coefs<-data.frame(sel_models3 %>% tidy(model))
sel_models3_coefs$sig<-ifelse(sel_models3_coefs$p.value<0.05,"*","") 
sel_models3_rsq<-data.frame(sel_models3 %>% glance(model))[1:3]
sel_models3_anova<-cbind(
  year=c("1987","1987","1987","1987","1987","1988","1988","1988","1988","1988",
         "1989","1989","1989","1989","1989","1990","1990","1990","1990","1990",
         "1991","1991","1991","1991","1991","1992","1992","1992","1992","1992",
         "1993","1993","1993","1993","1993","1994","1994","1994","1994","1994",
         "1995","1995","1995","1995","1995","1996","1996","1996","1996","1996",
         "2006","2006","2006","2006","2006","2007","2007","2007","2007","2007",
         "2008","2008","2008","2008","2008","2009","2009","2009","2009","2009",
         "2010","2010","2010","2010","2010","2011","2011","2011","2011","2011",
         "2012","2012","2012","2012","2012","2013","2013","2013","2013","2013",
         "2014","2014","2014","2014","2014","2015","2015","2015","2015","2015",
         "2016","2016","2016","2016","2016","2017","2017","2017","2017","2017"),
  variable=rep(c("FFD_std","I(FFD_std^2)","n_fl_std","I(n_fl_std^2)","FFD_std:n_fl_std"),22),
  plyr::ldply(lapply(as.list(sel_models3)$model,FUN=Anova), function(x) data.frame(x)[1:5,3:4]))
sel_models3_anova$sig<-ifelse(sel_models3_anova$Pr..F.<0.05,"*","") 

kable(sel_models3_coefs) #Selection gradients
#Correlational selection in 1988, 2016
#Quadratic selection on n_fl in 1990, 2006, 2015 and on FFD in 2015
kable(sel_models3_rsq) #R squares
kable(sel_models3_anova) #Anova results
```

Selection gradients for FFD for each year, from models including also number of flowers

```{r Selection gradients for each year, echo=FALSE}
ggplot(sel_grads_FFD,aes(estimate,year,colour=sig))+geom_point()+ylab("Selection gradient")+
  scale_color_manual(labels = c("Not significant", "Significant"), values = c("blue", "red"))+
  geom_vline(xintercept=0)+theme(legend.position="top")
```

# Models to explain variation in selection gradients among years

## How are yearly selection gradients linked to climatic variables?

```{r Effects of climatic variables on selection gradients}
sel_grads_FFD$signif<-as.factor(with(sel_grads_FFD,ifelse(sig=="*","1","0")))
sel_grads_FFD$sig<-NULL
names(sel_grads_FFD)<-c("year","selgradFFD","sig_selgradFFD")
data_sel_agg<-merge(mean_weather5,sel_grads_FFD)

#Fit univariate linear models of selgradFFD against each predictor (climatic variables)
varlist<-names(data_sel_agg)[c(8:10,20:22,32:34,42:46,56:58,68:70,80:82,
                 92:94,104:106,116:118,128:130,134,138:142,146:217,217,223:232)]
models_selgrads<-lapply(varlist, function(x) {
  summary(lm(substitute(selgradFFD ~ scale(i), list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_selgrads<-cbind(varlist,
              plyr::ldply(models_selgrads, function(x) coef(x)[2]),
              plyr::ldply(models_selgrads, function(x) coef(x)[8]),
              plyr::ldply(models_selgrads, function(x) x$adj.r.square))
names(models_selgrads)<-c("variable","estimate","p","adj.rsquare")
models_selgrads$sig<-ifelse(models_selgrads$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_selgrads,sig=="*"),desc(adj.rsquare)))
```

Only precipitation in March is significant. 

Plot of the best model

```{r Plots of best models, echo=FALSE, fig.height=3, fig.width=3.5}
  ggplot(data_sel_agg,aes(precipitation_3,selgradFFD))+geom_point()+geom_smooth(method="lm")+
  xlab("Sum of precipitation in March")+ylab("Selection gradient for FFD")
```

The model without interaction explains 18% of variance in selection gradients for FFD. Selection for early flowering (i.e. a more negative selection gradient) is stronger with lower precipitation in March. This means that when March is dry, there is more advantage of flowering early. Why?? The effect of precipitation in March might be related to the amount of snow and vegetative development? If March is dry (i.e. there is not a lot of snow), it is an advantage to start vegetative development earlier (which also would mean flowering earlier).

## How are yearly selection gradients linked to mean of FFD?

```{r Effects of FFD_mean on selection gradients}
kable(prettify(summary(lm(selgradFFD~FFD_mean,data=data_sel_agg)))) #No effect of FFD_mean
```

## How are yearly selection gradients linked to variance of FFD?

```{r Effects of FFD_var on selection gradients}
kable(prettify(summary(lm(selgradFFD~FFD_var,data=data_sel_agg)))) #No effect of FFD_var
```

## How are yearly selection gradients linked to range of FFD?

```{r Effects of FFD_dur on selection gradients}
kable(prettify(summary(lm(selgradFFD~FFD_dur,data=data_sel_agg)))) #No effect of FFD_dur
```

## How are yearly selection gradients linked to skewness of FFD?

```{r Effects of FFD_skew on selection gradients}
kable(prettify(summary(lm(selgradFFD~FFD_skew,data=data_sel_agg)))) #No effect of FFD_skew
```

## How are yearly selection gradients linked to kurtosis of FFD?

```{r Effects of FFD_kurt on selection gradients}
kable(prettify(summary(lm(selgradFFD~FFD_kurt,data=data_sel_agg)))) #No effect of FFD_kurt
```

## How are yearly selection gradients linked to intensity of grazing?

```{r Effects of grazing on selection gradients}
data_sel_agg<-merge(data_sel_agg,aggregate(grazing~year,data_sel,FUN=mean))

#Added mean grazing per year
kable(prettify(summary(lm(selgradFFD~grazing,data=data_sel_agg))))
      #p=0.0691 
kable(prettify(summary(lm(selgradFFD~grazing,data=subset(data_sel_agg,year<2017)))))
      #Removing 2017, worse p=0.0948 
#No effect of grazing (or very low effect?)
```

```{r Plots grazing and selection gradients, echo=FALSE, fig.height=3, fig.width=7}
grid.arrange(
  ggplot(data_sel_agg,aes(grazing,selgradFFD))+geom_point()+geom_smooth(method="lm")+
  xlab("Mean grazing")+ylab("Selection gradient for FFD")+ggtitle("With all years"),
ggplot(subset(data_sel_agg,year<2017),aes(grazing,selgradFFD))+geom_point()+geom_smooth(method="lm")+
  xlab("Mean grazing")+ylab("Selection gradient for FFD")+ggtitle("Without 2017"),ncol=2)
```

Although the effects are not significant, a higher grazing pressure seems to favor selection for later flowering.

## How are yearly selection gradients linked to seed predation?

```{r Effects of seed predation on selection gradients}
data_sel$n_pred_seeds<-with(data_sel,n_seeds-n_intact_seeds)
data_sel$prop_pred_seeds<-with(data_sel,
        ifelse(n_seeds==n_intact_seeds,0,
               (data_sel$n_seeds-data_sel$n_intact_seeds)/data_sel$n_seeds))
data_sel_agg<-merge(data_sel_agg,aggregate(prop_pred_seeds~year,data_sel,FUN=mean))
data_sel_agg<-merge(data_sel_agg,aggregate(n_pred_seeds~year,data_sel,FUN=mean))
data_sel_agg<-merge(data_sel_agg,aggregate(n_seeds~year,data_sel,FUN=mean))

#Added mean seed predation (proportion and n of predated seeds) per year
kable(prettify(summary(lm(selgradFFD~prop_pred_seeds,data=data_sel_agg)))) #NS
kable(prettify(summary(lm(selgradFFD~n_pred_seeds,data=data_sel_agg)))) #NS
```

No effect of seed predation on selection gradients.

## How are yearly selection gradients linked to fruit and seed set?

```{r Effects of fruit and seed set on selection gradients, fig.height=4, fig.width=8}
data_sel$fr_set<-with(data_sel,n_fr/n_fl)           #Calculate fruit set
data_sel$seed_set<-with(data_sel,n_seeds/n_ovules)  #Calculate seed set
nrow(subset(data_sel,n_fr>n_fl)) #0 cases where n_fruits>n_flowers
par(mfrow=c(1,2))
hist(data_sel$fr_set,main=NULL)
hist(data_sel$seed_set,main=NULL)
data_sel_agg<-merge(data_sel_agg,aggregate(fr_set~year,data_sel,FUN=mean))
data_sel_agg<-merge(data_sel_agg,aggregate(seed_set~year,data_sel,FUN=mean))

kable(prettify(summary(lm(selgradFFD~fr_set,data=data_sel_agg)))) #NS
kable(prettify(summary(lm(selgradFFD~seed_set,data=data_sel_agg)))) #NS
```

No effect of fruit and seed set on selection gradients.

## How are yearly selection gradients linked to number of seeds per flower?

```{r Effects of number of seeds per flower on selection gradients, fig.height=4, fig.width=8}
data_sel$seeds_per_fl<-with(data_sel,n_seeds/n_fl)  #Calculate n seeds per fl
hist(data_sel$seeds_per_fl,main=NULL)
data_sel_agg<-merge(data_sel_agg,aggregate(seeds_per_fl~year,data_sel,FUN=mean))

kable(prettify(summary(lm(selgradFFD~seeds_per_fl,data=data_sel_agg)))) #NS
```

No effect of number of seeds per flower on selection gradients.

## How are yearly selection gradients linked to grazing, seed predation, fruit and seed set? (alltogether)

```{r Effects of 4 predictors on selection gradients}
summary(lm(selgradFFD~grazing+prop_pred_seeds+fr_set+seed_set,data=data_sel_agg)) #Grazing*
summary(lm(selgradFFD~grazing+prop_pred_seeds+seeds_per_fl,data=data_sel_agg)) #NS
```
Grazing is significant: there is increased selection for late flowering in years of more intense grazing (as expected because grazers attack more the early-flowering ones).

# Models to explain variation in grazing and seed predation among years

```{r Variation in grazing and seed pred among years, echo=FALSE, fig.height=3, fig.width=9}
grid.arrange(
  ggplot(data_sel_agg,aes(year,grazing))+geom_point(),
  ggplot(data_sel_agg,aes(year,prop_pred_seeds))+geom_point(),
ncol=2)
```

## Grazing

### How is grazing linked to climatic variables?

```{r Effects of climatic variables on grazing}
#Fit univariate linear models of grazing against each predictor (climatic variables)
models_grazing<-lapply(varlist, function(x) {
  summary(lm(substitute(log(grazing) ~ scale(i), list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_grazing<-cbind(varlist,
              plyr::ldply(models_grazing, function(x) coef(x)[2]),
              plyr::ldply(models_grazing, function(x) coef(x)[8]),
              plyr::ldply(models_grazing, function(x) x$adj.r.squared)
              )

names(models_grazing)<-c("variable","estimate","p","rsquare")
models_grazing$sig<-ifelse(models_grazing$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_grazing,sig=="*"),desc(rsquare)))
```

It seems that grazing is higher when there is more precipitation in March and April-May are less warm. 

Model with two of the best variables: precipitation in March and GDD10 April+May

```{r Model grazing with the two best variables}
summary(lm(log(grazing)~scale(precipitation_3)+scale(GDD10_45),data=data_sel_agg))
```

This model explains 46% of the variation in grazing.

### How is grazing linked to mean of FFD?

```{r Effects of FFD_mean on grazing}
kable(prettify(summary(lm(log(grazing) ~ FFD_mean,data=data_sel_agg))))
      #No effect of FFD_mean 
kable(prettify(summary(lm(log(grazing) ~ FFD_mean,data=subset(data_sel_agg,!year=="2017")))))
#Removing 2017, No effect of FFD_mean 
```

### How is grazing linked to variance of FFD?

```{r Effects of FFD_var on grazing}
kable(prettify(summary(lm(log(grazing) ~ FFD_var,data=data_sel_agg))))
      #FFD_var*
summary(lm(log(grazing) ~ FFD_var,data=data_sel_agg))$adj.r.squared
kable(prettify(summary(lm(log(grazing) ~ FFD_var,data=subset(data_sel_agg,!year=="2017")))))
#Removing 2017, FFD_var*
summary(lm(log(grazing) ~ FFD_var,data=subset(data_sel_agg,!year=="2017")))$adj.r.squared
```

Grazing increases with variance of FFD.

### How is grazing linked to range of FFD?

```{r Effects of FFD_dur on grazing}
kable(prettify(summary(lm(log(grazing) ~ FFD_dur,data=data_sel_agg)))) #FFD_dur*
summary(lm(log(grazing) ~ FFD_dur,data=data_sel_agg))$adj.r.squared
kable(prettify(summary(lm(log(grazing) ~ FFD_dur,data=subset(data_sel_agg,!year=="2017"))))) 
summary(lm(log(grazing) ~ FFD_dur,data=data_sel_agg))$adj.r.squared
#Removing 2017, FFD_dur*
```

Grazing increases with range of FFD.

### How is grazing linked to skewness of FFD?

```{r Effects of FFD_skew on grazing}
kable(prettify(summary(lm(log(grazing) ~ FFD_skew,data=data_sel_agg))))
      #No effect of FFD_skew 
kable(prettify(summary(lm(log(grazing) ~ FFD_skew,data=subset(data_sel_agg,!year=="2017")))))
#Removing 2017, No effect of FFD_skew 
```

### How is grazing linked to kurtosis of FFD?

```{r Effects of FFD_kurt on grazing}
kable(prettify(summary(lm(log(grazing) ~ FFD_kurt,data=data_sel_agg))))
      #No effect of FFD_kurt 
kable(prettify(summary(lm(log(grazing) ~ FFD_kurt,data=subset(data_sel_agg,!year=="2017")))))
      #Removing 2017, No effect of FFD_kurt 
```

### Plots of best models grazing

```{r Plots of best models grazing, echo=FALSE, fig.height=6, fig.width=7}
grid.arrange(
  ggplot(data_sel_agg,aes(x=precipitation_3,y=log(grazing)))+geom_point()+geom_smooth(method="lm")+
  xlab("Sum of precipitation in March")+ylab("Grazing (log)"),
ggplot(data_sel_agg,aes(x=GDD10_45,y=log(grazing)))+geom_point()+geom_smooth(method="lm")+
  xlab("Sum of GDD10 in April and May")+ylab("Grazing (log)"),
ggplot(data_sel_agg,aes(x=FFD_var,y=log(grazing)))+geom_point()+geom_smooth(method="lm")+
  xlab("Variance of FFD")+ylab("Grazing (log)"),
ggplot(data_sel_agg,aes(x=FFD_dur,y=log(grazing)))+geom_point()+geom_smooth(method="lm")+
  xlab("Range of FFD")+ylab("Grazing (log)"),
ncol=2)
```

## Seed predation

### How is seed predation linked to climatic variables?

```{r Effects of climatic variables on seed predation}
#Fit univariate linear models of grazing against each predictor (climatic variables)
models_spred<-lapply(varlist, function(x) {
  summary(lm(substitute(asin(sqrt(prop_pred_seeds)) ~ scale(i), 
            list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_spred<-cbind(varlist,
              plyr::ldply(models_spred, function(x) coef(x)[2]),
              plyr::ldply(models_spred, function(x) coef(x)[8]),
              plyr::ldply(models_spred, function(x) x$adj.r.squared)
              )

names(models_spred)<-c("variable","estimate","p","rsquare")
models_spred$sig<-ifelse(models_spred$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_spred,sig=="*"),desc(rsquare)))
```

It seems that seed predation increases with temperature/GDD/GDH from May to June. 

Model with one of the best variables: 

```{r Model seed predation with one of the best variables}
summary(lm(asin(sqrt(prop_pred_seeds))~scale(max456),data=data_sel_agg))
```

This model explains 24% of the variation in seed predation

```{r Plots of best model seed predation, echo=FALSE, fig.height=3, fig.width=4}
ggplot(data_sel_agg,aes(x=max456,y=asin(sqrt(prop_pred_seeds))))+geom_point()+
  geom_smooth(method="lm")+xlab("Average maximum temperature April-June")+
  ylab("Proportion of predated seeds \n (arcsin-transformed)")
```

### How is seed predation linked to mean of FFD?

```{r Effects of FFD_mean on seed predation}
kable(prettify(summary(lm(sqrt(prop_pred_seeds) ~ FFD_mean,data=data_sel_agg)))) #No effect of FFD_mean
```

### How is seed predation linked to variance of FFD?

```{r Effects of FFD_var on seed predation}
kable(prettify(summary(lm(sqrt(prop_pred_seeds) ~ FFD_var,data=data_sel_agg)))) #No effect of FFD_var
```

### How is seed predation linked to range of FFD?

```{r Effects of FFD_dur on seed predation}
kable(prettify(summary(lm(sqrt(prop_pred_seeds) ~ FFD_dur,data=data_sel_agg)))) #No effect of FFD_dur
```

### How is seed predation linked to skewness of FFD?

```{r Effects of FFD_skew on seed predation}
kable(prettify(summary(lm(sqrt(prop_pred_seeds) ~ FFD_skew,data=data_sel_agg)))) #No effect of FFD_skew
```

### How is seed predation linked to kurtosis of FFD?

```{r Effects of FFD_kurt on seed predation}
kable(prettify(summary(lm(sqrt(prop_pred_seeds) ~ FFD_kurt,data=data_sel_agg)))) #No effect of FFD_kurt
```

# Models to explain variation in fruit and seed set among years

```{r Variation in fruit and seed set among years, echo=FALSE, fig.height=3, fig.width=9}
grid.arrange(
  ggplot(data_sel_agg,aes(year,fr_set))+geom_point(),
  ggplot(data_sel_agg,aes(year,seed_set))+geom_point(),
ncol=2)
```

## Fruit set

### How is fruit set linked to climatic variables?

```{r Effects of climatic variables on fruit set}
#Fit univariate linear models of fruit set against each predictor (climatic variables)
models_frset<-lapply(varlist, function(x) {
  summary(lm(substitute(fr_set ~ scale(i), list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_frset<-cbind(varlist,
              plyr::ldply(models_frset, function(x) coef(x)[2]),
              plyr::ldply(models_frset, function(x) coef(x)[8]),
              plyr::ldply(models_frset, function(x) x$adj.r.squared)
              )

names(models_frset)<-c("variable","estimate","p","rsquare")
models_frset$sig<-ifelse(models_frset$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_frset,sig=="*"),desc(rsquare)))
```

No significant relationships.

### How is fruit set linked to mean of FFD?

```{r Effects of FFD_mean on fruit set}
kable(prettify(summary(lm(fr_set ~ FFD_mean,data=data_sel_agg)))) #No effect of FFD_mean FFD 
```

### How is fruit set linked to variance of FFD?

```{r Effects of FFD_var on fruit set}
kable(prettify(summary(lm(fr_set ~ FFD_var,data=data_sel_agg)))) #FFD_var*
summary(lm(fr_set ~ FFD_var,data=data_sel_agg))$adj.r.square
```

Fruit set increases with variance of FFD.

```{r Plot effects of FFD_var on fruit set, echo=FALSE, fig.height=3, fig.width=3.5}
ggplot(data_sel_agg,aes(x=FFD_var,y=fr_set))+geom_point()+geom_smooth(method="lm")+
  xlab("Variance of FFD")+ylab("Fruit set")
```

### How is fruit set linked to range of FFD?

```{r Effects of FFD_dur on fruit set}
kable(prettify(summary(lm(fr_set ~ FFD_dur,data=data_sel_agg)))) #No effect of FFD_dur 
```

### How is fruit set linked to skewness of FFD?

```{r Effects of FFD_skew on fruit set}
kable(prettify(summary(lm(fr_set ~ FFD_skew,data=data_sel_agg)))) #No effect of FFD_skew  
```

### How is fruit set linked to kurtosis of FFD?

```{r Effects of FFD_kurt on fruit set}
kable(prettify(summary(lm(fr_set ~ FFD_kurt,data=data_sel_agg)))) #No effect of FFD_kurt  
```

##Seed set

### How is seed set linked to climatic variables?

```{r Effects of climatic variables on seed set}
#Fit univariate linear models of fruit set against each predictor (climatic variables)
models_seedset<-lapply(varlist, function(x) {
  summary(lm(substitute(seed_set ~ scale(i), list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_seedset<-cbind(varlist,
              plyr::ldply(models_seedset, function(x) coef(x)[2]),
              plyr::ldply(models_seedset, function(x) coef(x)[8]),
              plyr::ldply(models_seedset, function(x) x$adj.r.squared)
              )

names(models_seedset)<-c("variable","estimate","p","rsquare")
models_seedset$sig<-ifelse(models_seedset$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_seedset,sig=="*"),desc(rsquare)))
```

Model with one of the best variables: precipitation in June

```{r Model seed set with one of the best variables}
summary(lm(seed_set~scale(precipitation_6),data=data_sel_agg))
```

This model explains 28% of the variation in seed set

```{r Plots of best model seed set, echo=FALSE, fig.height=3, fig.width=4}
ggplot(data_sel_agg,aes(x=precipitation_6,y=seed_set))+geom_point()+
  geom_smooth(method="lm")+xlab("Sum of precipitation in June")+
  ylab("Seed set (n seeds/n ovules)")
```


### How is seed set linked to mean of FFD

```{r Effects of FFD_mean on seed set}
kable(prettify(summary(lm(seed_set ~ FFD_mean,data=data_sel_agg)))) #No effect of FFD_mean 
```

### How is seed set linked to variance of FFD

```{r Effects of FFD_var on seed set}
kable(prettify(summary(lm(seed_set ~ FFD_var,data=data_sel_agg)))) #No effect of FFD_var 
```

### How is seed set linked to range of FFD

```{r Effects of FFD_dur on seed set}
kable(prettify(summary(lm(seed_set ~ FFD_dur,data=data_sel_agg)))) #No effect of FFD_dur 
```

### How is seed set linked to skewness of FFD

```{r Effects of FFD_skew on seed set}
kable(prettify(summary(lm(seed_set ~ FFD_skew,data=data_sel_agg)))) #No effect of FFD_skew 
```

### How is seed set linked to kurtosis of FFD

```{r Effects of FFD_kurt on seed set}
kable(prettify(summary(lm(seed_set ~ FFD_kurt,data=data_sel_agg)))) #No effect of FFD_kurt 
```

# Models to explain variation in number of seeds per flower among years

## How is number of seeds per flower linked to climatic variables?

```{r Effects of climatic variables on seeds_per_fl, fig.height=3, fig.width=4}
hist(data_sel_agg$seeds_per_fl)
#Fit univariate linear models of fruit set against each predictor (climatic variables)
models_seeds_per_fl<-lapply(varlist, function(x) {
  summary(lm(substitute(seeds_per_fl ~ scale(i), list(i = as.name(x))),data=data_sel_agg))})

#Build a table with estimate, p and r square for all fitted models
models_seeds_per_fl<-cbind(varlist,
              plyr::ldply(models_seeds_per_fl, function(x) coef(x)[2]),
              plyr::ldply(models_seeds_per_fl, function(x) coef(x)[8]),
              plyr::ldply(models_seeds_per_fl, function(x) x$adj.r.squared)
              )

names(models_seeds_per_fl)<-c("variable","estimate","p","rsquare")
models_seeds_per_fl$sig<-ifelse(models_seeds_per_fl$p<0.05,"*","") # *=p<0.05

#Order models with significant variables by R square
kable(arrange(subset(models_seeds_per_fl,sig=="*"),desc(rsquare)))
```

No significant relationships.

## How is number of seeds per flower linked to mean of FFD

```{r Effects of FFD_mean on seeds_per_fl}
kable(prettify(summary(lm(seeds_per_fl ~ FFD_mean,data=data_sel_agg)))) #No effect of FFD_mean 
```

## How is number of seeds per flower linked to variance of FFD

```{r Effects of FFD_var on seeds_per_fl}
kable(prettify(summary(lm(seeds_per_fl ~ FFD_var,data=data_sel_agg)))) #FFD_var* 
summary(lm(seeds_per_fl ~ FFD_var,data=data_sel_agg))$adj.r.square
```

Number of seeds per flower increases with FFD_var (as did fruit set).

```{r Plot effects of FFD_var on seeds_per_fl, echo=FALSE, fig.height=3, fig.width=3.5}
ggplot(data_sel_agg,aes(x=FFD_var,y=seeds_per_fl))+geom_point()+geom_smooth(method="lm")+
  xlab("Variance of FFD")+ylab("Number of seeds per flower")
```

## How is number of seeds per flower linked to range of FFD

```{r Effects of FFD_dur on seeds_per_fl}
kable(prettify(summary(lm(seeds_per_fl ~ FFD_dur,data=data_sel_agg)))) #No effect of FFD_dur 
```

## How is number of seeds per flower linked to skewness of FFD

```{r Effects of FFD_skew on seeds_per_fl}
kable(prettify(summary(lm(seeds_per_fl ~ FFD_skew,data=data_sel_agg)))) #No effect of FFD_skew 
```

## How is number of seeds per flower linked to kurtosis of FFD

```{r Effects of FFD_kurt on seeds_per_fl}
kable(prettify(summary(lm(seeds_per_fl ~ FFD_kurt,data=data_sel_agg)))) #No effect of FFD_kurt 
```

# SUMMARY
* Phenotypic selection on flowering phenology differs among years (also when including number of flowers in the model)
* There is significant selection for early flowering on 7 out of 22 years studied
* Quadratic selection on flowering phenology is not important (only found in one year)
* Selection for early flowering decreases with precipitation in March and with grazing (although grazing is only significant when included as predictor in a model together with seed predation, fruit set and seed set)
* Grazing increases with precipitation in March and with variance and duration of flowering, and decreases with temperature in April-May
* Seed predation increases with temperature in April-June
* Fruit set increases with variance of flowering
* Seed set increases with precipitation in (April-May-)June
* The number of seeds per flower increases with variance of flowering



